---
title: "MATH5806 - Assignment 2 [100 marks]"
subtitle: '**Due Date**: 17:00, 13 October 2017'
output:
  pdf_document:
    keep_tex: yes
  html_document:
    keep_md: yes
  word_document: default
---


I declare that this assessment item is my own work, except where acknowledged, and has not been submitted for academic credit elsewhere. I acknowledge that the assessor of this item may, for the purpose of assessing this
item reproduce this assessment item and provide a copy to another member of the University; and/or communicate a copy of this assessment item to a plagiarism checking service (which may then retain a copy of the assessment
item on its database for the purpose of future plagiarism checking). I certify that I have read and understood the University Rules in respect of Student Academic Misconduct.

\bigskip
\bigskip

Name / Student Number: ... `r ( your_name <- "" )` / `r ( your_zID <- "" )` ...

Signature, Date: ... SIGN HERE THE PRINTED PDF VERSION ..., `r format(Sys.time(), "%d / %m / %Y")`

\bigskip
\bigskip

This document is an _R Markdown Notebook_ (http://rmarkdown.rstudio.com). To interact with this file, and to be able to produce a PDF file from it, you need to install three software on your computer (in that order):

1. **R:** https://cran.r-project.org/ and also `knitr`
2. **\LaTeX:** https://miktex.org/download
3. **RStudio:** https://www.rstudio.com/products/rstudio/download/

You can find more information about the _Markdown language_ here: 

- Tutorial: http://commonmark.org/help/tutorial/
- Try it online: https://upmath.me/
- Extensive guide: https://www.markdownguide.org/

and about the _R Markdown language_ here: 

- https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf 


In this R Notebook, you can add a new _chunk_ (i.e., a piece of R code that will be executable from within your document) by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*. When you execute code within the notebook, the results appear beneath the code. Each chunk can be executed by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. Here is an example:

```{r}
rnorm(5)
```


To create a PDF file containing the code and output, click the *Preview* button or press *Ctrl+Shift+K*.



It is easy to include mathematical formulas by using \LaTeX\ code. For inline equations, this code has to be put between single dollar signs, e.g., $\sum x_i$, while for centered equations double dollar signs must be used, e.g.,
$$
\int_0^1 f(x)dx =1.
$$

A nice introduction to \LaTeX\ can be found here: 

- https://www.sharelatex.com/blog/latex-guides/beginners-tutorial.html 

Finding individual symbols in \LaTeX\ is made easy via this online drawing tool:

- http://detexify.kirelabs.org/classify.html

```{r, echo = FALSE}
library("knitr")
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

-------------------------------------------------------------------------------------------------------------

\newpage

[**Instructions**]: # ( 1. Name your Rmd file as follows: StudentFirstName_StudentLastName_zID.Rmd (where zID is your zID)
                           For example: Pierre_LafayeDeMicheaux_z1234567.Rmd
                        2. Your Rmd code must compile (i.e., Knit to PDF) **without any error** (with `eval_mycode <- TRUE`, see below).
                        3. **Crucial:** Keep the names of the variables as I have defined them.
                        4. When you have finished, print the PDf file, sign it and bring it to me.
                      )


```{r, echo = TRUE}
# Put this variable to TRUE when you have completed all the codes below.
eval_mycode <- TRUE
```

# Exercise 1 : Multiple Choices Questions (MCQ) [25 marks]

Below, for A) to E), you need to complete the R chunks in order to:

- identify the **response variable** (give its CAPITALIZED NAME as provided in the text, e.g., `NAME`),
- give its **type** (one of "binary", "discrete", "continuous", "ordinal" or "qualitative"), 
- associate each data set with the most appropriate **regression technique** to use (one of "Simple Linear Regression", "Multiple Linear Regression", "ANOVA", "ANCOVA", "Logistic Regression", "Ordered Logistic Regression", "Poisson Regression" or "Multinomial Logit Model"), 
- give the **R function name** to perform that technique,
- give the **R package name** to which the function belongs.

Once you will have completed the R chunks, the corresponding sentences written below each chunk will be automatically filled. So you should **not** modify the sentences yourself. Also, do **not** leave white spaces within the quotation marks.

\bigskip

A) We analyze the factors that influence young adults leaving their parents' home through three pathways (`PATHWAY`): marriage, premarital residential independence, and residential semiautonomy in dormitories and barracks. Some other variables measured are the family income (`INC`), the gender (`GENDER`) and the number of years since high school (`YEAR`).
 
 
```{r, echo = TRUE}
# Replace the ... with your answer. Do NOT remove the "".
response_variable_name_A <- "PATHWAY"
type_var_A <- "qualitative"
regression_technique_A <- "MultinomialLogitModel"
R_function_name_A <- "multinom"
R_package_name_A <- "nnet"
```

_The response variable in data set A is ``r response_variable_name_A``. Its type is: `r type_var_A`. The regression technique to use to analyse this data set is called: `r regression_technique_A`. To perform it, one can use the ``r R_function_name_A`` function from the ``r R_package_name_A`` package._


B) Dioxin is a molecule that can cause cancer in humans and animals. We want to explain the variation of measured dioxin emission at a number of municipal solid waste (MSW) incinerators plants. Concentration of dioxin (in ng/$m^3$) were measured (`DIOXIN`) in 100 gas samples. The explanatory variables are:

- `PLANT`: the experiments have been conducted at three plants;
- `TEMP`: the temperature of each gas sample;
- `CO2`: the $CO_2$ concentration (in ppm) of each gas sample.

```{r, echo = TRUE}
# Replace the ... with your answer. Do NOT remove the "".
response_variable_name_B <- "DIOXIN"
type_var_B <- "continuous"
regression_technique_B <- "MultipleLinearRegression"
R_function_name_B <- "lm"
R_package_name_B <- "stats"
```

_The response variable in data set B is ``r response_variable_name_B``. Its type is: `r type_var_B`. The regression technique to use to analyse this data set is called: `r regression_technique_B`. To perform it, one can use the ``r R_function_name_B`` function from the ``r R_package_name_B`` package._

C) We want to predict if a person is likely to have an esophageal cancer (`ECANCER`). We consider a sample of 10,000 persons taken at random in the Australian population. The explanatory variables are the age of the person (`AGE`), as well as two binary indicators of smoking (`SMOKE`) and drinking (`DRINK`) habits.

```{r, echo = TRUE}
# Replace the ... with your answer. Do NOT remove the "".
response_variable_name_C <- "ECANCER"
type_var_C <- "binary"
regression_technique_C <- "LogisticRegression"
R_function_name_C <- "glm"
R_package_name_C <- "stats"
```

_The response variable in data set C is ``r response_variable_name_C``. Its type is: `r type_var_C`. The regression technique to use to analyse this data set is called: `r regression_technique_C`. To perform it, one can use the ``r R_function_name_C`` function from the ``r R_package_name_C`` package._

D) 50 fishes of the salmon species were caught in September 1997 at three specific locations (`LOC`) in the North sea and the number of parasites (`NBPAR`) were counted in each fish. We want to explain the number of parasites (`PARASITE`) as a function of the location and of the temperature (`TEMP`) of the sea.

```{r, echo = TRUE}
# Replace the ... with your answer. Do NOT remove the "".
response_variable_name_D <- "NBPAR"
type_var_D <- "discrete"
regression_technique_D <- "PoissonRegression"
R_function_name_D <- "glm"
R_package_name_D <- "stats"
```

_The response variable in data set D is ``r response_variable_name_D``. Its type is: `r type_var_D`. The regression technique to use to analyse this data set is called: `r regression_technique_D`. To perform it, one can use the ``r R_function_name_D`` function from the ``r R_package_name_D`` package._

E) We asked MATH5806 students if they think they have learnt something valuable during the course (variable `VALUE`). The possible answers are "Strongly Disagree", "Disagree", "Agree" and "Strongly Agree". We also record the age (`AGE`) of each student and if he/she has already worked (`WORKED`) in the private sector as a statistician or a data scientist.

```{r, echo = TRUE}
# Replace the ... with your answer. Do NOT remove the "".
response_variable_name_E <- "VALUE"
type_var_E <- "ordinal"
regression_technique_E <- "OrderedLogisticRegression"
R_function_name_E <- "polr"
R_package_name_E <- "nnet"
```

_The response variable in data set E is ``r response_variable_name_E``. Its type is: `r type_var_E`. The regression technique to use to analyse this data set is called: `r regression_technique_E`. To perform it, one can use the ``r R_function_name_E`` function from the ``r R_package_name_E`` package._



# Exercise 2 : Computer Simulation [36 marks]

In this exercise, you are going to explore the concept of **censoring**.

1. Generate independently $n=200$ observations $(x_i,y_i^*)$ from the model
$$
Y^* = \beta_1 + \beta_2 x + \epsilon,
$$
where $\beta_1=1$, $\beta_2=0.2$ and $\epsilon$ follows a standard normal distribution. The $x_i$'s are (randomly and independently) uniformly distributed between 0 and 14.
    ```{r, eval = TRUE}
    set.seed(1)
    n <- 200
    beta1 <- 1
    beta2 <- 0.2
    x <-  runif(n,0,14)
    eps <- rnorm(n)
    Ystar <- beta1+beta2*x+eps
    ```


2. Plot the simulated data, and add the estimated regression line (obtained by Simple Linear Regression).

    ```{r, eval = TRUE}
    # Do NOT change title, x- and y- labels, colours, etc
    plot(x,Ystar, main = "Regression without censoring", xlab = "x", ylab = "y*", col = "orangered", pch = 20, ylim = c(0, 8))
    abline(h = 2)
    ( SLR_lm <- lm(Ystar~x) )
    abline(SLR_lm, col = "blue")
        
    # Save plot for marking - Do not modify below
    my_par2.2 <- par()
    dev.copy(device = png, filename = 'pic2.2.png', width = 1000, height = 500) 
    dev.off()
    ```

3. Now, let us suppose that the following **censoring** occured: we do not observe (measure) the value of $y^*$ for any observation with a value of $y^*$ lower than some given threshold $\tau$. Instead of completely removing these cases (which would then be called **truncation**), we rather set the value of $y^*$ to some given value $\tau_y$. 

    This situation could happen for example at a charity reception. Every person present is obviously willing to make a donation of $y^*\$$. But if the minimum amount allowed for a donation is set to $\tau\$$, then some people might not be able to give something (and thus would give $\tau_y=0\$$). Consequently, censoring will occur, and we will rather only observe values of the following random variable:
$$
 Y_i=\left\{
\begin{array}{cl}
Y_i^* & \text{if }Y_i^*>\tau\\
\tau_y & \text{if }Y_i^*\leq \tau.
\end{array}
\right.
$$
     Now, let us get back to our simulated example. 
     
     Set $\tau=2$ and $\tau_y=2$. Plot the data $(x_i, y_i)$, $i=1,\ldots,n$ as above but use triangles (`pch = 17`) for the observations that have been censored (so that all observations below the horizontal line $h=2$ will be represented by triangles lying just above the $x$-axis). Add (\textcolor{blue}{in blue}) the fitted SLR line for the observations $(x_i, y_i^*)$ and (\textcolor{red}{in red}) the fitted SLR line for the observations $(x_i, y_i)$. Observe the discrepancy between the two lines. (Note that the aim would be to obtain the blue line only from the observations $(x_i, y_i)$, which is not done very well by the SLR modeling.)
     
     
    ```{r, eval = TRUE}
    tau <- 2
    tauy <- 2
    # Do NOT change title, x- and y- labels, colours, etc
    plot(x[which(Ystar>tau)],Ystar[which(Ystar>tau)], main = "Regression with censoring", xlab = "x", ylab = "y", col = "orangered", pch = 20, ylim = c(0, 8))
    points(x[which(Ystar<=tau)],rep(tauy,n)[which(Ystar<=tau)], pch = 17)
    abline(h = 2)
    abline(SLR_lm, col = "blue")
    Y <- Ystar ;  for(i in 1:n){
      if (Ystar[i]<=tau) Y[i]<-tauy}
    ( SLR_censored <- lm(Y~x) )
    abline(SLR_censored, col = "red")
    # Save plot for marking - Do not modify below
    my_par2.3 <- par()
    dev.copy(device = png, filename = 'pic2.3.png', width = 1000, height = 500) 
    dev.off()
    ```

4. We just need another modeling technique than the SLR. This technique is called the **Tobit model**. Find (Google search) an R function that can perform a Tobit regression and add (\textcolor{magenta}{in magenta}) the Tobit fitted line on the previous plot reproduced below. Observe the results to see if you were able to do better.

    ```{r, eval = TRUE}
    # Do NOT change title, x- and y- labels, colours, etc
    plot(x[which(Ystar>tau)],Ystar[which(Ystar>tau)], main = "Regression with Tobit model", xlab = "x", ylab = "y", col = "orangered", pch = 20, ylim = c(0, 8))
    points(x[which(Ystar<=tau)],rep(tauy,n)[which(Ystar<=tau)], pch = 17)
    abline(h = 2)
    abline(SLR_lm, col = "blue")
    abline(SLR_censored, col = "red")
    library("AER")
    ( tobit_model <- tobit(Ystar~x,left=tau) )
    abline(tobit_model, col = "magenta")
        
    # Save plot for marking - Do not modify below
    my_par2.4 <- par()
    dev.copy(device = png, filename = 'pic2.4.png', width = 1000, height = 500) 
    dev.off()
    ```
    
5. Let us compare the Residual Sums of Squares (RSS) for the two approaches (**computed with respect to the $y^*_i$ values**). Compute `RSS_lm` and `RSS_Tobit`, the RSS for the SLR and the Tobit approaches respectively. Comment the results.

    ```{r, eval = TRUE}
    ( RSS_SLR_censored <- sum(residuals(SLR_lm)^2) )
    ( RSS_Tobit <- sum(residuals(tobit_model)^2) )
    ```

> You comment here: The RSS for the SLR and the Tobit approaches are 196.9594 and 136.2085, respectively. The RSS for Tobit is smaller than that of SLR.

\newpage

# Exercise 3 : Analysis of a real data set [39 marks]

You are going to analyse the prestige of a scientist's first academic job as a university lecturer using the random variable `JOB` which can take values in the closed interval $[1;5]$. The prestige of the job was unavailable for scientists hired in a department where graduate programs were rated below 1.0 and for departments without graduate programs. The prestige of `JOB` for these cases have been recoded as 1, we are thus in presence of censoring. (We could also consider that we have a truncation and delete all these observations but we will not follow that road here.)

The table below describes all the variables and give a few descriptive statistics on them for $n=408$ scientists.


| Name   | Mean    | SD       | Minimum  | Maximum | Description                   |
|:-------|--------:|---------:|---------:|--------:|:------------------------------|
| JOB    | 2.23    | 0.97     | 1.00     | 4.80    | Prestige of job (from 1 to 5) |
| FEM    | 0.39    | 0.49     | 0.00     | 1.00    | 1 if female; 0 if male        |
| PHD    | 3.20    | 0.95     | 1.00     | 4.80    | Prestige of Ph.D. department  |
| MENT   | 45.47   | 65.53    | 0.00     | 532.0   | Citations received by mentor  |
| FEL    | 0.62    | 0.49     | 0.00     | 1.00    | 1 if held fellowship; else 0  |
| ART    | 2.28    | 2.26     | 0.00     | 18.00   | Number of articles published  |
| CIT    | 21.72   | 33.06    | 0.00     | 203.00  | Number of citations received  | 


1. Download the data set `tobjob2.xls` from Moodle. Copy and then Paste Special (Text and Numbers only) the data
in a new Excel (97/2003) file named `tobjob2_paste_special.xls`. Import these data into R.

    **Note:** If you cannot succeed using the `read.xls()` function on your system as suggested below, you can instead first export (from Excel) your file into a comma separated values (CSV) format, and then use the `read.csv()` function.

    ```{r, eval = TRUE}
    # This package enables one to read Excel files
    library("gdata")
    
    # Read the data set (do not use a PATH here)
    # Keep only the variables indicated in the above table
    prestige <- read.csv("tobjob2_paste_special.csv",header = TRUE)[,2:8]
    
    # We rename the columns to be exactly the same as in the Table above
    colnames(prestige) <- c("JOB", "FEM", "PHD", "MENT", "FEL", "ART", "CIT")

    # We display the first 10 lines
    head(prestige)
    ```
    
2.  Reproduce the numerical results of the above table.

    ```{r, eval = TRUE}
    # Sample size
    ( n_prestige <- nrow(prestige) )
    
    # Means
    ( means_prestige <- apply(prestige,2,'mean') )
      
    # Standard Deviations
    ( SDs_prestige <- apply(prestige,2,'sd') )
    
    # Minimums
    ( minimums_prestige <- apply(prestige,2,'min') )
    
    # Maximums
    ( maximums_prestige <- apply(prestige,2,'max') )
    ```
    
    
3. The Tobit model is defined through a **latent** random variable $Y^*$ that has been censored on the left, and whose expectation $\mathbb{E}(Y^*)$ is supposed to be linearly related to a set of predictors $\boldsymbol{x}=(x_1,\ldots,x_p)^T$:
$$
 Y_i=\left\{
\begin{array}{cl}
Y_i^*=\boldsymbol{x}_i^T\boldsymbol{\beta}+\epsilon_i & \text{if }\quad Y_i^*>\tau\\
\tau_y & \text{if }\quad Y_i^*\leq \tau,
\end{array}
\right.
$$
    with $\epsilon_i\sim\mathcal{N}(0,\sigma^2)$.
    
    What is a latent random variable? (Give a definition.)
    
    > Your response here: In statistics, latent random variables, are variables that are not directly observed. But They are rather inferred through a mathematical model from other variables that are observed.
    
    The threshold $\tau$ determines whether $Y^*$ is censored, while $\tau_y$ is the value assigned to $Y$ if $Y^*$ is censored. What are the values of $\tau$ and of $\tau_y$ for the Prestige data set:
    
    ```{r, eval = TRUE}
    ( tau_prestige <- 1 )
    ( tauy_prestige <- 1 )
    ```
    

4. Use an "official" R function as the one you provided in Exercise 2 to compute the estimates of the coefficients $\beta$'s and $\sigma^2$. 

    ```{r, eval = TRUE}
    library("AER")
    tobit_prestige <-  tobit(JOB~ .,left=tau_prestige,data=prestige)
    summary(tobit_prestige)
    ```
    
5. Interpret some of the results using the method of Partial Changes in $Y^*$.

> Your interpretation here: The most significant impact of these variables that PHD,FEL and CIT on the variable JOB. 

6. Denote respectively by $\Phi(\cdot)$ and $\phi(\cdot)$ the cumulative distribution function and the density function of the standard normal. Write the formula for the probability of "observation" $Y_i^*$ to being censored (as a function of $\Phi$, $\tau$, $\boldsymbol{x}_i$, $\boldsymbol{\beta}$ and $\sigma$):
    $$
    Pr(Y_i^*\leq \tau|\boldsymbol{x}_i) = Pr(\epsilon_i\leq\tau-\boldsymbol{x}_i^T\boldsymbol{\beta})=\Phi(\frac{\tau-
    \boldsymbol{x}_i^T\boldsymbol{\beta}}{\sigma})
    $$
    Write the log-likelihood (as a function of $y_i$, $\phi$, $\Phi$, $\tau$, $\boldsymbol{x}_i$, $\boldsymbol{\beta}$ and $\sigma$):
    $$
    \textrm{ln}~\mathcal{L}(\boldsymbol{\beta},\sigma^2|\boldsymbol{y},\boldsymbol{X}) = \sum_{\text{Uncensored}}\ln\phi(\frac{y_i-\boldsymbol{x}_i^T\boldsymbol{\beta}}{\sigma}) + \sum_{\text{Censored}} \ln\Phi(\frac{\tau-
    \boldsymbol{x}_i^T\boldsymbol{\beta}}{\sigma})
    $$
7. Create a function `loglik()` to compute the log-likelihood (**without using for loops**). Maximise the log-likelihood using the Prestige data set. Observe the value of the log-likelihood at the optimum and compare it with the one you found at question 4 above. Observe the values of the estimates for the coefficients $\beta$'s. Check that you find the same values than in question 4. You will use the R built-in optimisation function `optim()` with the "BFGS" method and the control argument `fnscale=-1`.

    ```{r, eval =TRUE}
    loglik <- function(theta, y, X, tau) {
      beta <- as.numeric(theta[1:7])
      sigma <- sqrt(abs(theta[8]))
      X <- as.matrix(X)
      y_i=y[which(y!=1)];
      x_i=X[which(y==1),];x_u=X[which(y!=1),]
      v=dnorm((y_i-x_u%*%beta)/sigma)/sigma
      c=1-pnorm((x_i%*%beta-tau)/sigma)
      Li=sum(log(v))+sum(log(c))
      return(Li)
    }
    y_prestige=prestige$JOB
    X_prestige=cbind(rep(1,n_prestige),prestige[,-1])
    starting_parameters=c(rep(0,7),1)
   ( optim_loglik <- optim(starting_parameters,loglik,method="BFGS",control=list(fnscale=-1),y=y_prestige,X=X_prestige,tau=tau_prestige))
    ```


